---
title: "BIM Resilience – PID HCR Tuning"
subtitle: "Generic PID-style Harvest Control Rule across ICES Category 1 Stocks"
author: "L.T. Kell"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: true
    number_sections: true
#bibliography: refs.bib
#biblio-style: apalike
vignette: >
  %\VignetteIndexEntry{Tuning a PID-style harvest control rule across ICES Category 1 stocks}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options:
  markdown:
    wrap: 72
---

# Introduction

This document implements a generic PID-style harvest control
rule (HCR) that adjusts TAC based on four potential indicators:
biomass state (`bstate`), fishing mortality state (`fstate`), biomass
slope (`btrnd`), and F slope (`ftrnd`). 

The evaluated for ICES Category 1 stocks within the FLR framework, found in the
`icesdata` package.

The aim is to find a default parameterisation (weights and gain terms)
that performs robustly across stocks, compare it to the ICES advice rule, and
then optimise in using management strategy
evaluation (MSE).

```{r knitr-setup, echo=FALSE}
knitr::opts_chunk$set(
  echo      =FALSE,
  warning   =FALSE,
  message   =FALSE,
  cache     =TRUE,
  cache.path="cache/13_pid_tuning/",
  fig.path  ="figs/13_pid_tuning/",
  dev       ="png",
  dpi       =300)

iFig =0
iTab =0
```


```{r setup}
library(FLCore)
library(FLBRP)
library(FLasher)
library(FLRebuild)
library(FLAssess)
library(FLife)
library(FLSRTMB)
library(FLSkill)

library(icesdata)
library(ggplot2)
library(ggplotFL)
library(plyr)
library(dplyr)
library(reshape)
library(RColorBrewer)
library(scales)
library(stringr)
library(ggpubr)
library(ggridges)

library(statcomp)
library(patchwork)

library(future.apply)
library(parallel)
library(doParallel)
library(foreach)

# For GA tuning (optional)
# install.packages("GA")
library(GA)
```

```{r}
source("C:/active/flr/FLSkill/R/slope.R")
source("C:/active/flrpapers/neaMac/source/boxFn.R")

source("C:/active/flr/FLSkill/R/generics.R")
source("C:/active/flr/FLSkill/R/roc.R")
source("C:/active/flr/FLSkill/R/classification.R")
source("C:/active/flr/FLSkill/R/utils.R")
source("C:/active/flr/FLSkill/R/plots.R")
```


# Data
```{r load-data}
data(icesdata)

ids=names(icesdata)
id=ids[35]

load("C:/active/flrpapers/haf/risk/data/lsmps.RData")

bbmsy=FLQuants(llply(icesdata, function(x) ssb( x)/benchmark(x)["btrigger"]))
ffmsy=FLQuants(llply(icesdata, function(x) fbar(x)/benchmark(x)["fmsy"]))

survey=FLQuants(llply(icesdata, function(x) rlnorm(1,log(tsb(x[-1])/mean(tsb(x[-1]))),0.2)))
hrate =FLQuants(llply(icesdata, function(x) rlnorm(1,log(catch(x)/ebiomass(x)),       0.2)))
lmn   =llply(lsmps, function(x) tryIt(lmean(x)))
```


```{r, biomass-state, fig.width=12,fig.height=6}
plot(mcf(FLQuants("B/B[MSY]"=bbmsy[[id]],"Survey"=survey[[id]])))+
  theme_minimal(16)
```

**Figure `r iFig=iFig+1; iFig`.** An example of comparison of an Operating Model quantity )spawning stock biomass (SSB) relative to the biomass at $B_{MSY}$) with an empricial indicator (total biomass less recruiting age class).

```{r, f-state, fig.width=12,fig.height=8}
plot(mcf(FLQuants("F/F[MSY]"=ffmsy[[id]],"Harvest~Rate"=hrate[[id]],"L[mean]"=1/lmn[[id]])))+
  theme_minimal(16)
```

**Figure `r iFig=iFig+1; iFig`.** An example of comparison of an Operating Model quantity (fishing mortality relative to  $F_{MSY}$) with empricial indicators (catch/exploitable biomass and inverse of the mean size).


# Skill

```{r, fig.width=12,fig.height=4}
dat=model.frame(FLQuants(response =ffmsy[[id]],
                         predictor=1/lmn[[id]]))
dat=transform(dat,response =response/median(response),
                  predictor=predictor/median(predictor))
ss=skillScore(response=dat$response,predictor=dat$predictor)
skillSummary( response=dat$response,predictor=dat$predictor)
skillPlot(    response=dat$response,predictor=dat$predictor)
```

**Figure `r iFig=iFig+1; iFig`.** 

```{r, roc, fig.width=12,fig.height=4}

biomass=mdply(ids, function(id) tryIt(cbind(.id=id,model.frame(FLQuants(bbmsy =bbmsy[[id]],
                                                                        survey=survey[[id]]),drop=TRUE))))[,-1]
f      =mdply(ids, function(id) tryIt(cbind(.id=id,model.frame(FLQuants(ffmsy =ffmsy[[id]],
                                                                        hrate =hrate[[id]],
                                                                        lmean=1/lmn[[id]]),drop=TRUE))))[,-1]

surveyROC=ddply(biomass,.(.id), with, roc2(response=bbmsy/mean(bbmsy),predictor=survey/mean(survey)))
 hrateROC=ddply(f,      .(.id), with, roc2(response=ffmsy/mean(ffmsy),predictor=hrate/mean(hrate)))
 lmeanROC=ddply(f,      .(.id), with, roc2(response=ffmsy/mean(ffmsy),predictor=lmean/mean(lmean)))

ggplot(rbind(cbind(What="Survey",      surveyROC),
             cbind(What="Harvest~Rate",hrateROC),
             cbind(What="L[mean]",     lmeanROC)))+
  geom_path(aes(FPR,TPR,group=.id))+
  geom_abline(aes(slope=1,intercept=0),col="red")+
  facet_wrap(~What)+
  theme_minimal(16)
```

**Figure `r iFig=iFig+1; iFig`.** 



```{r, State, fig.width=8,fig.height=4}
bstate=mdply(ids, function(id) 
  cbind(.id=id,model.frame(FLQuants(response=bbmsy[[id]],predictor=survey[[id]]),drop=TRUE)))[,-1]
bstate=ddply(bstate,.(.id), with, tryIt(skillSummary(response=response,predictor=predictor/mean(predictor))))

fstate=mdply(ids, function(id) 
  cbind(.id=id,model.frame(FLQuants(response=ffmsy[[id]],predictor=hrate[[id]]),drop=TRUE)))[,-1]
fstate=ddply(fstate,.(.id), with, tryIt(skillSummary(response=response,predictor=predictor/mean(predictor))))

lstate=mdply(names(lmn[laply(lmn, is.FLQuant)]), function(id) 
  cbind(.id=id,model.frame(mcf(FLQuants(response=ffmsy[[id]],predictor=1/lmn[[id]])),drop=TRUE)))[,-1]
lstate=ddply(lstate,.(.id), with, tryIt(skillSummary(response=response,predictor=predictor/mean(predictor))))

state=rbind.fill(cbind(Indicator="Survey",                 bstate),
                 cbind(Indicator="Relative\nHarvest Rate", fstate),
                 cbind(Indicator="Length-Based\nIndicator",lstate))

ggplot(state, aes(x=AUC, y=Indicator)) +
  geom_density_ridges(
    stat="binline",
    bins=20,
    scale=0.9,
    draw_baseline=FALSE) +
  xlab("AUC") +
  ylab("")+
  theme_minimal(16)
```

**Figure `r iFig=iFig+1; iFig`.** 

```{r slope, fig.width=8,fig.height=4}
btrnd=mdply(ids, function(id) { 
  tryIt(cbind(.id=id,model.frame(FLQuants(response =slope(bbmsy[[id]]/mean(bbmsy[[id]])),
                                          predictor=slope(survey[[id]]/mean(survey[[id]]))),drop=TRUE)))})[,-1]
btrnd=ddply(btrnd,.(.id), with, tryIt(skillSummary(response=exp(response),predictor=exp(predictor))))

ftrnd=mdply(ids, function(id) 
  tryIt(cbind(.id=id,model.frame(FLQuants(response=slope(ffmsy[[id]]/mean(ffmsy[[id]])),
                                          predictor=slope(hrate[[id]]/mean(hrate[[id]]))),drop=TRUE))))[,-1]
ftrnd=ddply(ftrnd,.(.id), with, tryIt(skillSummary(response=exp(response),predictor=exp(predictor))))

ltrnd=mdply(names(lmn[laply(lmn, is.FLQuant)]), function(id) 
  tryIt(cbind(.id=id,model.frame(mcf(FLQuants(response=slope(ffmsy[[id]]/mean(ffmsy[[id]])),
                                              predictor=slope(1/lmn[[id]]*mean(lmn[[id]])))),drop=TRUE))))[,-1]
ltrnd=ddply(ltrnd,.(.id), with, tryIt(skillSummary(response=exp(response),predictor=exp(predictor))))

trnd=rbind.fill(cbind(Indicator="Survey",                 btrnd),
                cbind(Indicator="Relative\nHarvest Rate", ftrnd),
                cbind(Indicator="Length-Based\nIndicator",ltrnd))

ggplot(trnd, aes(x=AUC, y=Indicator)) +
  geom_density_ridges(
    stat="binline",
    bins=20,
    scale=0.9,
    draw_baseline=FALSE
  ) +
  xlab("AUC") +
  ylab("")+
  theme_minimal(16)
```

**Figure `r iFig=iFig+1; iFig`.** 

```{r, entropy, fig.width=12,fig.height=4}
ent=ldply(bbmsy, function(x) tryIt(data.frame(entropy=permutation_entropy(ordinal_pattern_distribution(x=x, ndemb=5)))))

ggplot(ent, aes(x=entropy,y=1)) +
  geom_density_ridges(
    stat="binline",
    bins=20,
    scale=0.9,
    draw_baseline=FALSE
  ) +
  xlab("Entropy") +
  ylab("")+
  theme_minimal(16)
```
**Figure `r iFig=iFig+1; iFig`.** 

```{r, biomass, , fig.width=12,fig.height=8}
hist=rbind(cbind(What="BB[MSY]",ldply(bbmsy, as.data.frame,drop=TRUE)),
           cbind(What="FF[MSY]",ldply(ffmsy, as.data.frame,drop=TRUE)))

fn<-function(x) {
  r=quantile(x, probs=c(0.05, 0.33, 0.5, 0.66, 0.66))
  names(r)=c("ymin", "lower", "middle", "upper", "ymax")
  r}

dt =ddply(hist,.(What,.id), subset, year==max(year))
dtB=subset(dt,What=="BB[MSY]")

ggplot(hist,aes(x=factor(.id,levels=dtB$.id[order(dtB$data)]),y=data))+
  facet_grid(What~.)+
  stat_summary(fun.data=fn, geom="boxplot")+
  geom_point(aes(x=.id,y=data),data=dt,col="red")+
  geom_hline(aes(yintercept=1),col="red")+
  theme_minimal(16)+
  theme(legend.position="bottom",strip.text.y=element_text(angle=90),
                                 strip.text.x=element_text(angle=0),
                                 axis.text.x=element_text(angle=45, hjust=1))+
  xlab("Stock")+ylab(expression(B/B[MSY]))
```

**Figure `r iFig=iFig+1; iFig`.** 


# PID-style HCR: nextCatch function

A generic P-only HCR on log-catch, with AUC-based weights and entropy-based damping.

## Design rationale

The rule operates on four error terms (biomass state, F state, biomass
trend, F trend) weighted by their diagnostic quality (AUC - Area Under
ROC Curve). The composite error drives a proportional control on
log-catch. High entropy (signal uncertainty) damps the response to
avoid chasing noise.

## Implementation

```{r define-nextCatch}
```



# Indicator construction

# PID HCR within an MSE loop

```{r test, evl=FALSE}
#' Run MSE with PID-style HCR for a single stock
#'
#' @param om    FLStock operating model.
#' @param eq    FLBRP equilibrium reference points.
#' @param start Integer. First projection year.
#' @param end   Integer. Last projection year.
#' @param nits  Integer (default 100). Number of iterations/replicates.
#' @param Kp Numeric (default 0.5). Proportional gain.
#' @param gMin Numeric (default 0.2). Minimum entropy-based gain.
#' @param maxChange Numeric (default 0.3). Max ± annual TAC change.
#' @param lambda* Numeric (default 1.0 each). Weight multipliers for indicators.
#'
#' @return FLStock with projected catch guided by PID HCR.
#'
#' @details
#'   For each year in [start, end), computes indicators (bstate, fstate,
#'   btrnd, ftrnd, entropy, AUC) and calls nextCatch() to determine
#'   next-year TAC. Projects forward using fwd() with recruitment
#'   deviations sampled from the stock-recruit residuals.
pidMSE<-function(om,eq, 
                 auc=rep(1,4),ent=1, 
                 start=dims(om)$maxyear, end=start+30, 
                 nits =dims(om)$iter,
                 sd=0.3, b=0.5,
                 n           =3,
                 Kp          =0.5,
                 gMin        =0.2,
                 maxChange   =0.3,
                 lambdabstate=1,
                 lambdafstate=1,
                 lambdaBtrnd =1,
                 lambdaFtrnd =1) {

  # Prepare OM window for projection
  om =fwdWindow(om, eq, end=end)

  # Extract and prepare recruitment deviations
  sr =attributes(eq)$sr
  recDevs =FLQuant(
    exp(sample(c(residuals(sr) - median(residuals(sr))),
               (end - start + 1) * nits, replace=TRUE)),
    dimnames=list(year=start:end, iter=seq(nits)))

  # Optional: (can be turned off if not needed)
  shortCut =rlnoise(nits, FLQuant(0, dimnames=list(year=start:end)),sd=sd, b=b)

  ## Data 
  getMSY<-function(x){
    fbar(x)=FLQuant(refpts(x)["msy","harvest",drop=T])
    
    tsbMSY=quantSums(stock.wt(x)[-1]%*%stock.n(x)[-1])
    MSY   =catch(x)
    hMSY  =catch(x)%/%ebiomass(x)
    ssb   =ssb(x)
    f     =fbar(x)
    
    rbind(FLPar("tsb"=tsbMSY[drop=TRUE]),FLPar("yield"=MSY[drop=TRUE]),FLPar("hrate"=hMSY[drop=TRUE]),
          FLPar("ssb"=ssb[drop=TRUE]),FLPar("harvest"=f[drop=TRUE]))}
  
  msy=getMSY(eq)
  
  MP=FLQuants(catch =catch(om),
              survey=tsb(om[-1])%/%msy["tsb"],
              hrate =(catch(om)%/%ebiomass(om[-1]))%*%msy["hrate"],
              ssb   =ssb( om[-1])%/%msy["ssb"],
              f     =fbar(om[-1])%/%msy["harvest"])

  MP[["sTrnd"]]=window(slope(MP$survey[,ac((start-n):start)],n=n),end=end)
  MP[["hTrnd"]]=window(slope(MP$hrate[, ac((start-n):start)],n=n),end=end)
  MP[["ssb"]]  =window(slope(MP$ssb[,   ac((start-n):start)],n=n),end=end)
  MP[["f"]]    =window(slope(MP$f[,     ac((start-n):start)],n=n),end=end)
                
  MP$catch[ ,ac(start:end)]=as.numeric(NA)
  MP$survey[,ac(start:end)]=as.numeric(NA)
  MP$hrate[ ,ac(start:end)]=as.numeric(NA)
  MP$ssb[   ,ac(start:end)]=as.numeric(NA)
  MP$f[     ,ac(start:end)]=as.numeric(NA)

  # Iterate over projection years applying PID HCR
  for (iYr in start:(end)) {

    MP$catch[,ac(iYr)]=catch(om)[,ac(iYr-1)]
    
if (!FALSE){

    MP$survey[,ac(iYr-1)]=tsb(om[-1])[,ac(iYr-1)]%/%msy["tsb"]
    MP$hrate[ ,ac(iYr-1)]=(MP$catch[,ac(iYr-1)]%/%ebiomass(om)[,ac(iYr-1)])%/%msy["harvest"]
    
    MP$sTrnd[ ,ac(iYr-1)]=trend(MP$survey[,ac(iYr-(n:1))],n=n)
    MP$hTrnd[ ,ac(iYr-1)]=trend(MP$hrate[ ,ac(iYr-(n:1))],n=n)
    
    for (it in seq_len(nits)) {
      
      MP$catch[,ac(iYr),,,,it] =
      nextCatch(
        bstate       =MP$survey[,ac(iYr-1),,,,it],
        fstate       =MP$hrate[,ac(iYr-1),,,,it],
        btrnd        =MP$sTrnd[,ac(iYr-1),,,,it],
        ftrnd        =MP$hTrnd[,ac(iYr-1),,,,it],
        aucBstate    =auc[1],
        aucFstate    =auc[2],
        aucBtrnd     =auc[3],
        aucFtrnd     =auc[4],
        ent          =ent,
        CNow         =c(MP$catch[,ac(iYr-1),,,,it]),
        lambdaBstate =lambdaBstate,
        lambdaFstate =lambdaFstate,
        lambdaBtrnd  =lambdaBtrnd,
        lambdaFtrnd  =lambdaFtrnd,
        Kp           =Kp,
        entLow       =0,
        entHigh      =1,
        gMin         =gMin,
        maxChange    =maxChange)
      
      }}
    
    om =FLasher::fwd(om,control=fwdControl(year =iYr,
                                           quant="catch",
                                           value=MP$catch[,ac(iYr)]),
              sr       =eq,
              deviances=recDevs[, ac(iYr)])}
#MP<<-FLQuants(llply(MP,window, start=start+1))
  om}
```


```{r eqs}
sgs=FLBRPs(llply(icesdata,eql,model="segreg"))  

R0 =ldply(sgs, function(x) data.frame(r0=c(params(x)["a"]*params(x)["b"])))

bhs=foreach(id=ids, .packages=c("FLCore","FLBRP","FLCandy")) %dopar% {
  eql(icesdata[[id]], model="bevholtSV", prior_r0=R0[R0$.id == id,"r0"], cv_r0=0.1)}
bhs=FLBRPs(bhs)
names(bhs)=ids 
```


```{r, PID,  fig.width=10,fig.height=5}
save(bhs,file="C:/tmp/bhs.RData")

rtn=pidMSE(propagate(icesdata[[27]],100),bhs[[27]],Kp=0.5,
           lambdaBstate =1.0,
           lambdaFstate =1.0,
           lambdaBtrnd  =0.5,
           lambdaFtrnd  =0.5)

plot(rtn)+
  geom_vline(aes(xintercept=2021),col="blue")+
  theme_minimal(16)
```
**Figure `r iFig=iFig+1; iFig`.** 


```{r}
fwdBIM<-function(stk,eql,ctc=NULL,nits=100,fmsy=benchmark(stk)["fmsy"],
               maxYear=dims(stk)$maxyear+30,f.cv=0.2,bnd=c(0.85,1.2),seed=1234){
        set.seed(seed)
  
        discards.n( stk)[is.na(discards.n(stk))] =1e-12
        discards.wt(stk)[is.na(discards.wt(stk))]=1e-12
  
        currentYear=dims(stk)$maxyear
        stk    =fwdWindow(stk,end=maxYear)
          
        if (is.na(fmsy)) return(NULL)
        
        recDevs=attributes(attributes(eql)$sr)$residuals
        recDevs=recDevs-median(recDevs)
        
        recDevs=FLQuant(sample(c(recDevs),dim(recDevs)[2]*nits,TRUE),
                        dimnames=list(year=currentYear:maxYear,iter=seq(nits)))

        fmsy   =FLQuant(c(fmsy),dimnames=list(year=currentYear:maxYear))
        fmsy   =rlnorm(nits,log(fmsy),f.cv)

        if (!is.null(ctc) && currentYear<dims(ctc)$maxyear){
          
          stk=FLasher::fwd(stk,catch=propagate(ctc[,ac(currentYear:dims(ctc)$maxyear)],nits),
                      sr   =eql,deviances=exp(recDevs))
          }
          
        if (any(is.na(bnd)))
          stk=FLCore::ffwd(stk,f=fmsy,sr=eql,deviances=exp(recDevs))
        else{
          for (iYr in ac((currentYear):maxYear)){
            stk=FLCore::ffwd(stk,f=fmsy[,iYr],sr=eql,deviances=exp(recDevs))
            ctc=qmin(qmax(catch(stk)[,iYr],catch(stk)[,ac(an(iYr)-1)]*bnd[1]),catch(stk)[,ac(an(iYr)-1)]*bnd[2])
            stk=FLasher::fwd(stk,catch=ctc,sr=eql,deviances=exp(recDevs))}}
        
        stk}   

rtn2=fwdBIM(icesdata[[27]],bhs[[27]])

plot(rtn2)
```

```{r}
plot(FLStocks("PID"=rtn,"ICES~AR"=rtn2)) 
```


We embed `nextCatch()` in a forward-projection MSE similar to the ICES
HCR example in your existing code.

```{r pid-mse-wrapper}
#' Run MSE with PID-style HCR for a single stock
#'
#' @param om    FLStock operating model.
#' @param eq    FLBRP equilibrium reference points.
#' @param start Integer. First projection year.
#' @param end   Integer. Last projection year.
#' @param nits  Integer (default 100). Number of iterations/replicates.
#' @param Kp Numeric (default 0.5). Proportional gain.
#' @param gMin Numeric (default 0.2). Minimum entropy-based gain.
#' @param maxChange Numeric (default 0.3). Max ± annual TAC change.
#' @param lambda* Numeric (default 1.0 each). Weight multipliers for indicators.
#'
#' @return FLStock with projected catch guided by PID HCR.
#'
#' @details
#'   For each year in [start, end), computes indicators (bstate, fstate,
#'   btrnd, ftrnd, entropy, AUC) and calls nextCatch() to determine
#'   next-year TAC. Projects forward using fwd() with recruitment
#'   deviations sampled from the stock-recruit residuals.
pidMSE<-function(om,eq,
                 start, end, nits=1,
                 Kp       =0.5,
                 gMin     =0.2,
                 maxChange=0.3,
                 lambdabstate=1.0,
                 lambdafstate=1.0,
                 lambdaBtrnd =1.0,
                 lambdaFtrnd =1.0) {

  # Prepare OM window for projection
  om =fwdWindow(om, eq, end=end)

  # Extract and prepare recruitment deviations
  sr =attributes(eq)$sr
  recDevs =FLQuant(
    exp(sample(c(residuals(sr) - median(residuals(sr))),
               (end - start + 1) * nits, replace=TRUE)),
    dimnames=list(year=start:end, iter=seq(nits))
  )

  # Optional: observation error (can be turned off if not needed)
  shortCut =rlnoise(nits,
                      FLQuant(0, dimnames=list(year=start:end)),
                      sd=0.3, b=0.5)

  # Iterate over projection years applying PID HCR
  for (iYr in start:(end - 1)) {

    # Extract current year's catch
    CNow =quantSums(catch(om)[, ac(iYr)])

    # Subset OM data up to current year for indicator calculations
    omHist =om[, ac(dims(om)$minyear:iYr)]

    # Calculate state (bstate, fstate) at current year
    st     =calcState(omHist, eq)

    # Calculate trends
    tr    =calcTrend(window(omHist,end=iYr-1), n=3)
   
    # Calculate entropy from biomass series
    B_series =quantSums(ssb(omHist))
    ent =calcEntropy(B_series)

    # Get AUC values (precomputed or estimated here)
    aucs =calcAUC(omHist, eq)

    # Apply nextCatch() separately per iteration
    CNext =numeric(nits)

    for (it in seq_len(nits)) {
      CNext[it] =nextCatch(
        bstate       =st$bstate,
        fstate       =st$fstate,
        btrnd        =tr$btrnd,
        ftrnd        =tr$ftrnd,
        aucBstate    =aucs["aucBstate"],
        aucFstate    =aucs["aucFstate"],
        aucBtrnd     =aucs["aucBtrnd"],
        aucFtrnd     =aucs["aucFtrnd"],
        ent          =ent,
        CNow         =c(CNow)[it],
        lambdaBstate =lambdaBstate,
        lambdaFstate =lambdaFstate,
        lambdaBtrnd  =lambdaBtrnd,
        lambdaFtrnd  =lambdaFtrnd,
        Kp           =Kp,
        entLow       =0,
        entHigh      =1,
        gMin         =gMin,
        maxChange    =maxChange
      )
    }

    # Forward projection: apply PID-determined catch
    om =FLasher::fwd(om,
              control=fwdControl(
                year =iYr + 1,
                quant="catch",
                value=CNext[1]
              ),
              sr      =eq,
              deviances= recDevs[, ac(iYr + 1)])
  }

  om}
```

# Multi-stock MSE and performance metrics

We run the PID HCR across all Category 1 stocks in parallel.



```{r pid-mse-all, eval=FALSE}
start =2020
end   =2030
nits  =100

# Example PID parameters (these will be tuned)
Kp        =0.4
gMin      =0.2
maxChange =0.3

lambdaBstate =1.0
lambdaFstate =1.0
lambdaBtrnd  =1.0
lambdaFtrnd  =1.0


ncores =detectCores() - 2
cl     =makeCluster(ncores)
registerDoParallel(cl)

msePID =foreach(id=ids[35],Kp=0.5,
                  .packages=c("FLCore", "FLBRP", "FLasher", "icesdata")) %dopar% {
  try({
     pidMSE(om=icesdata[[id]], 
            eq=bhs[[id]],
            start=start, end=end, nits=nits,
            Kp=Kp, gMin=gMin, maxChange=maxChange,
            lambdabstate=lambdaBstate,
            lambdafstate=lambdaFstate,
            lambdaBtrnd =lambdaBtrnd,
            lambdaFtrnd =lambdaFtrnd)
  }, silent=TRUE)}

stopCluster(cl)
names(msePID) =ids

# Drop failed runs
msePID =msePID[!sapply(msePID, inherits, "try-error")]

# Save results
save(msePID, file="C:/active/bim/tex/report/data/results/msePID.RData")
```

# Visualising PID TAC trajectories

```{r plot-pid-tac, fig.width=10, fig.height=12,eval=FALSE}
ctcPID =FLQuants(mlply(names(msePID), function(id)
  catch(msePID[[id]]) / attributes(msePID[[id]])$eqsim[["Catchequi"]]
))

names(ctcPID) =names(msePID)
names(ctcPID)[grep("her.27.1-24a514a", names(ctcPID))] ="her.27.1-24"

plot(ctcPID) +
  geom_path(
    aes(year, data), col="grey80",
    data=subset(ldply(names(ctcPID),
                        function(qname)
                          cbind(qname=qname,
                                as.data.frame(iter(ctcPID[[qname]], 1),
                                              drop=TRUE))),
                  year > 2024)
  ) +
  geom_hline(aes(yintercept=1), col="grey") +
  facet_wrap(~qname, scale="free_y", ncol=4) +
  labs(
    title="Cat 1: PID HCR – Catch / MY",
    x="Year",
    y=expression(Catch/MY)
  ) +
  theme_minimal(base_size=12) +
  theme(
    legend.position  ="bottom",
    legend.title     =element_blank(),
    axis.text        =element_blank(),
    axis.ticks       =element_blank(),
    panel.grid.minor =element_blank(),
    panel.grid.major =element_line(color="grey90"),
    plot.title       =element_text(face="bold", size=14),
    plot.subtitle    =element_text(size=12, color="grey30"),
    panel.spacing    =unit(0, "lines")
  )
```

**Figure `r iFig =iFig + 1; iFig`.** TACs under the PID HCR (catch
relative to MY) for Category 1 stocks.

# Performance metrics and risk evaluation

```{r compute-performance-metrics,eval=FALSE}
#' Compute summary metrics across stocks and iterations
#'
#' @param mseResults List of FLStock projections from MSE.
#' @param start Integer. First projection year.
#' @param end Integer. Last projection year.
#' @param Blim Numeric. Biomass limit reference point.
#'
#' @return Data frame with one row per stock, columns for yield, risk, IAV.
compute_metrics<-function(mseResults, start, end, Blim=NULL) {

  metrics =ldply(names(mseResults), function(id) {
    stk =mseResults[[id]]

    # Mean catch over projection
    meanCatch =mean(c(catch(stk)[, ac(start:end)]), na.rm=TRUE)

    # Interannual variability in catch
    ctc_vec =c(catch(stk)[, ac(start:end)])
    iav =sd(diff(ctc_vec)) / mean(ctc_vec, na.rm=TRUE)

    # Risk: proportion of years/iters with SSB below limit
    ssb_vec =c(ssb(stk)[, ac(start:end)])
    if (is.null(Blim)) {
      Blim =0.25 * median(ssb_vec, na.rm=TRUE)
    }
    risk_Blim =mean(ssb_vec < Blim, na.rm=TRUE)

    data.frame(
      StockID  =id,
      MeanCatch=meanCatch,
      IAV      =iav,
      Risk_Blim=risk_Blim
    )
  })

  metrics
}

# Compute metrics for current parameter set
perf =compute_metrics(msePID, start=start, end=end)

print(perf)
```

```{r summary-table,eval=FALSE}
iTab =iTab + 1

summary_table =perf %>%
  summarise(
    Stocks=n(),
    Mean_Catch=mean(MeanCatch, na.rm=TRUE),
    SD_Catch  =sd(MeanCatch, na.rm=TRUE),
    Mean_IAV  =mean(IAV, na.rm=TRUE),
    Max_Risk  =max(Risk_Blim, na.rm=TRUE),
    Stocks_Below_Risk=sum(Risk_Blim > 0.05)
  )

knitr::kable(summary_table,
             caption=paste0("Table ", iTab, ". Summary performance metrics across stocks"))
```

# Parameter tuning with a genetic algorithm

A genetic algorithm can efficiently search the parameter space to find
parameter sets that meet risk constraints and maximise yield.

```{r ga-fitness-function,eval=FALSE}
#' Fitness function for GA tuning
#'
#' @param theta Numeric vector of parameters:
#'   theta[1]=Kp
#'   theta[2]=gMin
#'   theta[3]=maxChange
#'   theta[4]=lambdaBstate
#'   theta[5]=lambdaFstate
#'   theta[6]=lambdaBtrnd
#'   theta[7]=lambdaFtrnd
#'
#' @return Numeric fitness (higher is better).
#'
#' @details
#'   Runs a (reduced) MSE with the parameter set, computes performance
#'   metrics across stocks, and returns a fitness combining yield and
#'   risk. The function penalises parameter sets that violate risk
#'   constraints (e.g. max risk > 0.05).
pid_fitness<-function(theta) {

  Kp        =theta[1]
  gMin      =theta[2]
  maxChange =theta[3]
  lambdaB   =theta[4]
  lambdaF   =theta[5]
  lambdaDB  =theta[6]
  lambaDF  =theta[7]

  # Guard against invalid parameter ranges
  if (Kp < 0 || gMin < 0 || gMin > 1 || maxChange < 0 || maxChange > 1) {
    return(-1e6)
  }

  # Run a reduced MSE for speed (subset of stocks or fewer iterations)
  subSIDs =ids  # or: ids[1:min(10, length(ids))]
  nits_ga =50        # fewer iters than full MSE for speed

  results =foreach(id=subSIDs,
                     .packages=c("FLCore", "FLBRP", "FLasher", "icesdata")) %dopar% {
    try({
      om =icedata[[id]]
      eq =eql(om, model="bevholtSV")

      pidMSE(om, eq,
                  start=start, end=end, nits=nits_ga,
                  Kp=Kp, gMin=gMin, maxChange=maxChange,
                  lambdabstate=lambdaB,
                  lambdafstate=lambdaF,
                  lambdaBtrnd =lambdaDB,
                  lambdaFtrnd =lambdaDF)
    }, silent=TRUE)
  }

  # Compute metrics
  valid =!sapply(results, inherits, "try-error")
  results =results[valid]

  if (length(results) < length(subSIDs) / 2) {
    # Many failures; penalise
    return(-1e6)
  }

  perf_ga =compute_metrics(results, start=start, end=end)

  # Aggregate fitness
  yield_mean =mean(perf_ga$MeanCatch, na.rm=TRUE)
  risk_max   =max(perf_ga$Risk_Blim, na.rm=TRUE)
  iav_mean   =mean(perf_ga$IAV, na.rm=TRUE)

  # Objective: maximise yield, penalise risk above 0.05, minimise IAV
  R0    =0.05
  alpha =100
  beta  =1

  fitness =yield_mean - alpha * max(0, risk_max - R0) - beta * iav_mean

  return(fitness)
}
```

```{r ga-tuning, eval=FALSE}
# Run GA for parameter tuning
# (Set eval=FALSE} only if you have significant compute time available)

# Set up parallel backend for GA
ncores_ga =detectCores() - 2
cl_ga     =makeCluster(ncores_ga)
registerDoParallel(cl_ga)

# Run GA
ga_res =ga(
  type="real-valued",
  fitness=pid_fitness,
  lower=c(
    0.05,   # Kp
    0.1,    # gMin
    0.1,    # maxChange
    0.5,    # lambdaBstate
    0.5,    # lambdaFstate
    0.5,    # lambdaBtrnd
    0.5     # lambdaFtrnd
  ),
  upper=c(
    1.0,    # Kp
    0.6,    # gMin
    0.4,    # maxChange
    3.0,    # lambdaBstate
    3.0,    # lambdaFstate
    3.0,    # lambdaBtrnd
    3.0     # lambdaFtrnd
  ),
  popSize =20,
  maxiter =30,
  parallel=TRUE
)

stopCluster(cl_ga)

# Extract best solution
best_params =ga_res@solution
names(best_params) =c("Kp", "gMin", "maxChange",
                        "lambdaBstate", "lambdaFstate",
                        "lambdaBtrnd", "lambdaFtrnd")

print(best_params)
print(summary(ga_res))

# Save GA results
save(ga_res, best_params,
     file="C:/active/bim/tex/report/data/results/ga_tuning_results.RData")
```

# Alternative: Grid search tuning

If GA is not available or you prefer explicit grid search, use this
approach.

```{r grid-search, eval=FALSE}
# Define parameter grid
param_grid =expand.grid(
  Kp        =c(0.2, 0.4, 0.6),
  gMin      =c(0.2, 0.3),
  maxChange =c(0.2, 0.3),
  lambdabstate=c(1.0, 1.5),
  lambdafstate=c(1.0, 1.5),
  lambdaBtrnd =c(0.8, 1.0),
  lambdaFtrnd =c(0.8, 1.0)
)

nrow(param_grid)  # Total number of parameter combinations

# Run grid search with progress reporting
results_grid =foreach(i=seq_len(nrow(param_grid)),
                        .combine=rbind) %dopar% {

  theta =unlist(param_grid[i, ])
  fitness =pid_fitness(theta)

  cbind(param_grid[i, ], fitness=fitness)
}

# Sort by fitness (descending)
results_grid =results_grid[order(results_grid$fitness, decreasing=TRUE), ]

# Display top 10 parameter sets
print(head(results_grid, 10))

# Save grid search results
save(results_grid, param_grid,
     file="C:/active/bim/tex/report/data/results/grid_search_results.RData")
```

# Visualising tuning results

Once GA or grid search completes, plot fitness landscape.

```{r plot-ga-fitness, eval=FALSE}
# If using GA results:
plot(ga_res)

# If using grid search results:
ggplot(results_grid, aes(x=Kp, y=gMin, fill=fitness)) +
  geom_tile() +
  facet_wrap(~lambdaBstate) +
  scale_fill_gradient2(low="red", high="green") +
  labs(title="Grid search fitness landscape",
       x="Kp", y="gMin", fill="Fitness") +
  theme_minimal()
```

# Final validation with best parameters

Once you identify optimal parameters (from GA or grid search), re-run the
full MSE with those parameters and visualise results.

```{r final-validation, eval=FALSE}
# Use best parameters found during tuning
Kp_final        =best_params["Kp"]
gMin_final      =best_params["gMin"]
maxChange_final =best_params["maxChange"]
lambdaBstate_final =best_params["lambdaBstate"]
lambdaFstate_final =best_params["lambdaFstate"]
lambdaBtrnd_final  =best_params["lambdaBtrnd"]
lambdaFtrnd_final  =best_params["lambdaFtrnd"]

# Re-run MSE with full number of iterations (e.g. nits=500)
msePID_final =foreach(id=ids,
                        .packages=c("FLCore", "FLBRP", "FLasher", "icesdata")) %dopar% {
  try({
    om =icedata[[id]]
    eq =eql(om, model="bevholtSV")

    pidMSE(om=om, eq=eq,
                start=start, end=end, nits=500,
                Kp=Kp_final,
                gMin=gMin_final,
                maxChange=maxChange_final,
                lambdabstate=lambdaBstate_final,
                lambdafstate=lambdaFstate_final,
                lambdaBtrnd =lambdaBtrnd_final,
                lambdaFtrnd =lambdaFtrnd_final)
  })
}

names(msePID_final) =ids
msePID_final =msePID_final[!sapply(msePID_final, inherits, "try-error")]

# Compute final performance
perf_final =compute_metrics(msePID_final, start=start, end=end)
print(perf_final)

# Save final results
save(msePID_final, perf_final,
     file="C:/active/bim/tex/report/data/results/msePID_final.RData")
```

# Summary and conclusions

This template provides:

1. **`nextCatch()` function**: A generic, stock-agnostic PID-style HCR
   that uses four indicators weighted by AUC and damped by entropy.

2. **Indicator helpers**: Placeholder functions for computing bstate,
   fstate, trends, entropy, and AUC values from FLStock objects.

3. **MSE framework**: Full integration into an FLR-based forward
   projection loop, with parallel evaluation across stocks.

4. **Tuning methods**: Both genetic algorithm and grid-search approaches
   to find global parameters that perform robustly across 82 stocks.

You can adapt this by:

- Replacing placeholder indicator functions with your actual
  implementations.
- Adjusting risk thresholds, fitness objectives, and parameter bounds
  to reflect your management goals.
- Running GA or grid search to find optimal parameters.
- Validating the best parameterisation on held-out stocks or longer
  time horizons.

# References
